---
title: "LAMForPublication"
author: "Saisakul CHernbumroong"
date: "8 October 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages
```{r}
library(tidyverse)
library(caret)
library(missForest)
library(ggplot2)
library(rJava)
library(RWeka)
library(FSelector)
library(mlbench)
library(clusterCrit)
library(MDAS)
library(mice)
library(FactoMineR)
library(dplyr)
library(gridExtra)
library(dplyr)
library(Hmisc)
library(xtable)
library(factoextra)
library(ConsensusClusterPlus)
library(cluster)
```


## Load data set
```{r}
#1) Load data
raw = read.csv("LAM Key variables edited data August 17.csv", header = TRUE, na.strings = c("#N/A","") ,strip.white = TRUE)
dim(raw)
summary(raw)

datatype = read.csv("Data type.csv", header = TRUE,strip.white = TRUE)
dim(datatype)
summary(datatype)

#Check variables' names
MDAS::checkVariableNames(raw, datatype$Parameter)


# Reformat data 
raw$SOB[is.na(raw$SOB)] = "0"
raw$PTX.[is.na(raw$PTX.)] = "0"
raw$other.resp[is.na(raw$other.resp)] = "0"
raw$angiomyolipoma[is.na(raw$angiomyolipoma)] = "0"
raw$other.non.resp[is.na(raw$other.non.resp)] = "0"
raw$Screened[is.na(raw$Screened)] = "0"
raw$None[is.na(raw$None)] = "0"

raw$TSC[is.na(raw$TSC)] = "0"
raw$AML.[is.na(raw$AML.)] = "0"
raw$LYMPHATIC[is.na(raw$LYMPHATIC)] = "0"
raw$pneumothorax[is.na(raw$pneumothorax)] = "0"

#2) Check data validity
#processed data to the correct type
valid = correctType(raw, datatype$Type)

checkValidity(valid, datatype$Type, showOnlyInvalid = F, saveToFile = T, 
              fileName = paste("valid", as.character(Sys.Date()), sep = "-"))

#Recode change 1-Yes, 0-No
factor.class = which(sapply(valid, class) == "factor")
for(col in factor.class)
{
  if(all(levels(valid[, col]) == c("0","1")))
  {
    valid[, col] = factor(ifelse(valid[, col] == "0", "No", ifelse(valid[, col] == "1", "Yes", NA)))
  }
}

#for menopause only
valid$menopause = as.character(valid$menopause)
valid$menopause = factor(ifelse(valid[, "menopause"] == "No", 
                                "Pre", "Post"))

checkValidity(valid, datatype$Type, showOnlyInvalid = F, saveToFile = T, 
              fileName = paste("valid", as.character(Sys.Date()), sep = "-"))

#Only use data with definite outcome in the analysis
#Dat is the data set with only defiite LAM outcomes
dat = valid[valid$definite...probable == "d", ] 

#Remove study code and definite outcome columns
dat = dat[, -c(1:2)]
```

## Create patient characteristics summary
```{r}
col.id = which(sapply(dat, is.numeric))
summary_table = matrix(0, nrow = 0, ncol = 7)
for(i in col.id)
{
  summary_table = rbind.data.frame(summary_table, summary(dat[, i]))
  summary_table[nrow(summary_table), 7] = sum(is.na(dat[, i]))
}

summary_table = cbind.data.frame("Variable" = colnames(dat)[col.id], summary_table)
colnames(summary_table) = c("Variable", names(summary(dat[, i])))

print(xtable(summary_table, caption = "Patients characteristics"), include.rownames = FALSE)

```

## Analyse missing data
```{r}
mis = missingAnalysis(dat)
missing.values <- dat %>%
  gather(key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  group_by(key) %>%
  mutate(total = n()) %>%
  group_by(key, total, isna) %>%
  summarise(num.isna = n()) %>%
  mutate(pct = num.isna / total * 100)


levels <-
  (missing.values  %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
  ggplot() +
  geom_bar(aes(x = reorder(key, desc(pct)), 
               y = pct, fill=isna), 
           stat = 'identity', alpha=0.8) +
  scale_x_discrete(limits = levels) +
  scale_fill_manual(name = "", 
                    values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
  coord_flip() +
  labs(title = "Percentage of missing values", x =
         'Variable', y = "% of missing values")

percentage.plot
```


## Investigate imputation techniques i.e. MICE+RF, RF, MICE
```{r}
seedno = 1574
set.seed(seedno)

mi.rf = mice(dat[, 1:25], defaultMethod = c("rf", "rf", "rf", "rf"), seed = seedno)

MDAS::checkImpute(dat[, 1:25], mice::complete(mi.rf))

im = missForest(dat[, 1:25])
MDAS::checkImpute(dat[, 1:25], im$ximp)

mi = mice(dat[, 1:25], seed = seedno)
MDAS::checkImpute(dat[, 1:25], mice::complete(mi))
```

## Identify number of clusters

```{r}
set.seed(seedno)
all.ca = c()
ca.ca = c()
for(i in 1:5)
{
  com = mice::complete(mi,i)
  
  res = FAMD(com[, 1:25], ncp = Inf, graph = F)
  
  ca = MDAS::clusterAnalysis(res$ind$coord, plot = F, repeated = 5)
  ca.ca = rbind(ca.ca, ca$clust)
  all.ca = rbind(all.ca,ca$best)
}
re = all.ca %>% group_by(Method) %>% summarise(mode(No.cluster, "all")[1])
re
write.csv(ca.ca, file = "IdentClust.csv")
write.csv(all.ca, file = "IdentClustMajority.csv")

all.ca = c()
ca.ca = c()
for(i in 1:5)
{
  com = mice::complete(mi,i)
  
  dt = daisy(com)
  
  ca = MDAS::clusterAnalysis(dt, diss = T, plot = F, repeated = 5)
  ca.ca = rbind(ca.ca, ca$clust)
  all.ca = rbind(all.ca,ca$best)
}
re = all.ca %>% group_by(Method) %>% summarise(mode(No.cluster, "all")[1])
re

# save results
write.csv(ca.ca, file = "IdentClustDist.csv")
write.csv(all.ca, file = "IdentClustDistMajority.csv")
```

## Here we look at the PCA results
```{r}
fviz_contrib(res, choice = "var", axes = 1, top = 10)
fviz_screeplot(res, addlabels = TRUE, ylim = c(0, 5, 0))
```


## Identify number of cluster using Gap statistics with Kmean
```{r}
for(i in 1:5)
{
  com = mice::complete(mi,i)
  
  res = FAMD(com[, 1:25], ncp = Inf, graph = F)
  plot(fviz_nbclust(res$ind$coord, kmeans, method = "gap"))
}
```

## Identify number of cluster using Gap statistics with PAM
```{r}
for(i in 1:5)
{
  com = mice::complete(mi,i)
  
  res = FAMD(com[, 1:25], ncp = Inf, graph = F)
  plot(fviz_nbclust(res$ind$coord, pam, method = "gap"))
}
```

## Identify number of cluster using Gap statistics with Hierarchical + Kmean
```{r}
for(i in 1:5)
{
  com = mice::complete(mi,i)
  
  res = FAMD(com[, 1:25], ncp = Inf, graph = F)
  plot(fviz_nbclust(res$ind$coord, hClustCon, method = "gap"))
}
```

## Identify number of cluster using Gap statistics with Fuzzy C-mean
```{r}
for(i in 1:5)
{
  com = mice::complete(mi,i)
  
  res = FAMD(com[, 1:25], ncp = Inf, graph = F)
  plot(fviz_nbclust(res$ind$coord, cmeans, method = "gap"))
}

```

## Perform cluster analysis using the identified number of clusters
```{r}
com = mice::complete(mi,1)
all.ca = clusterPlotAnalysis(com, k = 2, plot = T, use.PCA = T)
for(i in 2:5)
{
  com = mice::complete(mi,i)
  
  cpa = MDAS::clusterPlotAnalysis(com, k = 2, plot = T, use.PCA = T)
  all.ca = cbind(all.ca,cpa)
}

#save results
write.csv(all.ca, file = "AllPatientscluster.csv")


```



```{r}
#Here we reorder the cluster names as each clustering algorithm randomly assigns the cluster names.
a= all.ca
a = a[, -which(colnames(a) == "EM")]

a$Kmeans = as.character(a$Kmeans)
a$Kmeans[a$Kmeans == 2] = 22
a$Kmeans[a$Kmeans == 1] = 2
a$Kmeans[a$Kmeans == 22] = 1
a$Kmeans = factor(a$Kmeans)

a$Fuzzy = as.character(a$Fuzzy)
a$Fuzzy[a$Fuzzy == 2] = 22
a$Fuzzy[a$Fuzzy == 1] = 2
a$Fuzzy[a$Fuzzy == 22] = 1
a$Fuzzy = factor(a$Fuzzy)

a$Kmeans.1 = as.character(a$Kmeans.1)
a$Kmeans.1[a$Kmeans.1 == 2] = 22
a$Kmeans.1[a$Kmeans.1 == 1] = 2
a$Kmeans.1[a$Kmeans.1 == 22] = 1
a$Kmeans.1 = factor(a$Kmeans.1)

a$Fuzzy.1 = as.character(a$Fuzzy.1)
a$Fuzzy.1[a$Fuzzy.1 == 2] = 22
a$Fuzzy.1[a$Fuzzy.1 == 1] = 2
a$Fuzzy.1[a$Fuzzy.1 == 22] = 1
a$Fuzzy.1 = factor(a$Fuzzy.1)

a$HCPC.2 = as.character(a$HCPC.2)
a$HCPC.2[a$HCPC.2 == 2] = 22
a$HCPC.2[a$HCPC.2 == 1] = 2
a$HCPC.2[a$HCPC.2 == 22] = 1
a$HCPC.2 = factor(a$HCPC.2)

a$`H-Kmeans.2`= as.character(a$`H-Kmeans.2`)
a$`H-Kmeans.2`[a$`H-Kmeans.2` == 2] = 22
a$`H-Kmeans.2`[a$`H-Kmeans.2` == 1] = 2
a$`H-Kmeans.2`[a$`H-Kmeans.2` == 22] = 1
a$`H-Kmeans.2` = factor(a$`H-Kmeans.2`)

a$`H-Fuzzy.2` = as.character(a$`H-Fuzzy.2`)
a$`H-Fuzzy.2`[a$`H-Fuzzy.2` == 2] = 22
a$`H-Fuzzy.2`[a$`H-Fuzzy.2` == 1] = 2
a$`H-Fuzzy.2`[a$`H-Fuzzy.2` == 22] = 1
a$`H-Fuzzy.2` = factor(a$`H-Fuzzy.2`)

a$Kmeans.2 = as.character(a$Kmeans.2)
a$Kmeans.2[a$Kmeans.2 == 2] = 22
a$Kmeans.2[a$Kmeans.2 == 1] = 2
a$Kmeans.2[a$Kmeans.2 == 22] = 1
a$Kmeans.2 = factor(a$Kmeans.2)

a$Hierarchical.2 = as.character(a$Hierarchical.2)
a$Hierarchical.2[a$Hierarchical.2 == 2] = 22
a$Hierarchical.2[a$Hierarchical.2 == 1] = 2
a$Hierarchical.2[a$Hierarchical.2 == 22] = 1
a$Hierarchical.2 = factor(a$Hierarchical.2)

a$HCPC.3 = as.character(a$HCPC.3)
a$HCPC.3[a$HCPC.3 == 2] = 22
a$HCPC.3[a$HCPC.3 == 1] = 2
a$HCPC.3[a$HCPC.3 == 22] = 1
a$HCPC.3 = factor(a$HCPC.3)

a$HCPC.4 = as.character(a$HCPC.4)
a$HCPC.4[a$HCPC.4 == 2] = 22
a$HCPC.4[a$HCPC.4 == 1] = 2
a$HCPC.4[a$HCPC.4 == 22] = 1
a$HCPC.4 = factor(a$HCPC.4)

a$Fuzzy.4 = as.character(a$Fuzzy.4)
a$Fuzzy.4[a$Fuzzy.4 == 2] = 22
a$Fuzzy.4[a$Fuzzy.4 == 1] = 2
a$Fuzzy.4[a$Fuzzy.4 == 22] = 1
a$Fuzzy.4 = factor(a$Fuzzy.4)

final.cluster = factor(apply(a, 1, mode))
```

#Get cluster descriptions and profile plots on imputed data sets.
```{r}
for(i in 1:5)
{
  com = mice::complete(mi,i)
  plotClusterProfile(com, final.cluster)
  getClusterDesc(com[!is.na(final.cluster),], final.cluster[!is.na(final.cluster)])
}

```


```{r}
#save results
g = list()
for(i in 1:5)
{
  com = mice::complete(mi,i)
  g[[i]] = plotHeatmapGroup(com[, sapply(com, is.numeric)], final.cluster) + ggtitle(paste("Imputation set", i))
}
do.call(grid.arrange, g)
for(i in 1:5)
{
com = mice::complete(mi,i)
cd = getClusterDesc(com[!is.na(final.cluster),], final.cluster[!is.na(final.cluster)])
write.csv(cd$numerical$cluster.1, file = paste0("numericalclusterdesc1_im", i,".csv"))
write.csv(cd$numerical$cluster.2, file = paste0("numericalclusterdesc2_im", i,".csv"))
write.csv(cd$categorical$cluster.1, file = paste0("categoricalclusterdesc1_im", i,".csv"))
write.csv(cd$categorical$cluster.2, file = paste0("categoricalclusterdesc2_im", i,".csv"))
write.csv(cd$desc, file = paste0("overallclusterdesc_im", i, ".csv"))
}
```

## Perform biomarker identification
```{r}
fss = list()
for(i in 1:5)
{
  com = mice::complete(mi,i)
  fs = fsAnalysis(com[!is.na(final.cluster),], final.cluster[!is.na(final.cluster)])
  fss[[i]] = fs  
}

```

```{r}

#Analyse feature selection results

re.fs = c()

for(i in 1:5)
{
  
  fs1 = fss[[i]]
    
  d = data.frame("Imputation" = i,
                 "Method" = "REF+NB",
             "No.Of.Selelected.Feature" = fs1$RFE$bestSubset,
             "Accuracy" = fs1$RFE$results$Accuracy[fs1$RFE$bestSubset],
             "AccuracySD" = fs1$RFE$results$AccuracySD[fs1$RFE$bestSubset],
             "Kappa" = fs1$RFE$results$Kappa[fs1$RFE$bestSubset],
             "KappaSD" = fs1$RFE$results$KappaSD[fs1$RFE$bestSubset])
  
  e = data.frame("Imputation" = i,
                 "Method" = "CFS+NB",
             "No.Of.Selelected.Feature" = fs1$CFS$no.selectedFeature,
             "Accuracy" = fs1$CFS$results$Accuracy[fs1$CFS$no.selectedFeature],
             "AccuracySD" = fs1$CFS$results$AccuracySD[fs1$CFS$no.selectedFeature],
             "Kappa" = fs1$CFS$results$Kappa[fs1$CFS$no.selectedFeature],
             "KappaSD" = fs1$CFS$results$KappaSD[fs1$CFS$no.selectedFeature])
  
  f = data.frame("Imputation" = i,
                 "Method" = "MRMR+NB",
             "No.Of.Selelected.Feature" = fs1$MRMR$no.selectedFeature,
             "Accuracy" = fs1$MRMR$results$Accuracy[fs1$MRMR$no.selectedFeature],
             "AccuracySD" = fs1$MRMR$results$AccuracySD[fs1$MRMR$no.selectedFeature],
             "Kappa" = fs1$MRMR$results$Kappa[fs1$MRMR$no.selectedFeature],
             "KappaSD" = fs1$MRMR$results$KappaSD[fs1$MRMR$no.selectedFeature])
  
  re.fs = rbind.data.frame(re.fs, d, e, f) 
  
}

```


## Analyse and combine biomarker selected from RFE in imputed data sets
```{r}

re.s = c()

for(i in 1:5)
{
  
  fs1 = fss[[i]]
    
  re.s = c(re.s, fs1$RFE$optVariables)
}

d = data.frame(summary(factor(re.s)))
d = cbind.data.frame(rownames(d), d)
colnames(d) = c("Variable", "Count")

d = d[order(d$Count, decreasing = T),]

sl = as.character(d[d$Count == 5, "Variable"])
save(sl, file = "selectedfeature")

```


## Perform machine learning analysis 
```{r}
#combine data sets
comdf = c()
outdf = c()

for(i in 1:5)
{
  comdf = rbind.data.frame(comdf, mice::complete(mi, i))
  outdf = c(outdf, final.cluster)
}

df = cbind.data.frame(comdf, outdf)

df = df[complete.cases(df), ]

df$outdf = factor(paste0("Cluster", df$outdf))

ml = mlAnalysis(df[, sl], df$outdf, nfold = 5, nrun = 10)

#save models
save(ml, file = "MLwithmissingFS")

```


```{r}
#model results
allm = rbind.data.frame(sapply(mm[mm$metric == "LogLoss",], mean),
                 sapply(mm[mm$metric == "AUC",], mean),
                 sapply(mm[mm$metric == "prAUC",], mean),
                 sapply(mm[mm$metric == "Accuracy",], mean),
                 sapply(mm[mm$metric == "Kappa",], mean),
                 sapply(mm[mm$metric == "F1",], mean),
                 sapply(mm[mm$metric == "Sensitivity",], mean),
                 sapply(mm[mm$metric == "Specificity",], mean),
                 sapply(mm[mm$metric == "Pos_Pred_Value",], mean),
                 sapply(mm[mm$metric == "Neg_Pred_Value",], mean),
                 sapply(mm[mm$metric == "Precision",], mean),
                 sapply(mm[mm$metric == "Recall",], mean),
                 sapply(mm[mm$metric == "Detection_Rate",], mean),
                 sapply(mm[mm$metric == "Balanced_Accuracy",], mean))
                 

```


